|

### 2021-09-22 paper.Rmd
date

---
title: |
  | Emotional framing and the effectiveness of 
  | corrective information^[I thank Thomas Bräuninger, Sebastian Juhl, the GIP team, and participants of the Fall 2020 CDSS workshop in political science at the University of Mannheim for excellent comments and helpful suggestions on this study. This work was supported by the German Research Foundation (DFG) via the Collaborative Research Center (SFB) 884 on "The Political Economy of Reforms" (Project C2) and the University of Mannheim’s Graduate School of Economics and Social Sciences.]
author: |
  | *work in progress*
  |
  | Pirmin Stöckle^[Corresponding address: pirmin.stoeckle@gess.uni-mannheim.de]
  | GESS and SFB 884, University of Mannheim
date: September 16, 2021
linestretch: 1.5
colorlinks: true
abstract: \setstretch{1} \noindent Concerns about various forms of misinformation and its fast dissemination through online media have generated huge interest into ways to effectively correct false claims. An under-explored mechanism in this research is the role of distinct emotions. How do emotional appeals interact with corrective information? Specifically, I focus on the emotion of disgust, which has been shown to be linked to the moralization of attitudes, which in turn reduces the impact of empirical evidence on attitudes and makes compromise less likely. Substantively, I investigate the issue of genetically modified (GM) food. I implement a pre-registered survey experiment manipulating both misinformation and corrective information treatments within a panel study based on a probability sample of the general population in Germany. Results show, first, that an emotionally framed piece of misinformation can increase moralization, replicating previous findings. Second, a neutral correction to the misinformation is only effective if the misnformation does not include a strong emotional frame, whereas a correction embedded in a counterframe is effective for both low and high emotional misinformation. In a broader perspective, these results inform further studies and policy interventions on other issues where disinformation loads on strong emotions, ranging from social policy over immigration to health interventions such as vaccinations. \vspace{.8cm}
# abstract-template: Everyone agrees that this issue is really important. But we do not know much about this specific question, although it matters a great deal, for these reasons. We approach the problem from this perspective. Our research design focuses on these cases and relies on these data, which we analyze using this method. Results show what we have learned about the question. They have these broader implications.

>>>>

---
title: |
  | Emotional framing and the effectiveness of 
  | corrective information^[I thank Thomas Bräuninger, Sebastian Juhl, the GIP team, and participants of the Fall 2020 CDSS workshop in political science at the University of Mannheim for excellent comments and helpful suggestions on this study. This work was supported by the German Research Foundation (DFG) via the Collaborative Research Center (SFB) 884 on "The Political Economy of Reforms" (Project C2) and the University of Mannheim’s Graduate School of Economics and Social Sciences.]
author: |
  | *work in progress*
  |
  | Pirmin Stöckle^[Corresponding address: pirmin.stoeckle@gess.uni-mannheim.de]
  | GESS and SFB 884, University of Mannheim
date: \today
linestretch: 1.5
colorlinks: true
abstract: \setstretch{1} \noindent Concerns about various forms of misinformation and its fast dissemination through online media have generated huge interest into ways to effectively correct false claims. An under-explored mechanism in this research is the role of distinct emotions. How do emotional appeals interact with corrective information? Specifically, I focus on the emotion of disgust, which has been shown to be linked to the moralization of attitudes, which in turn reduces the impact of empirical evidence on attitudes and makes compromise less likely. Substantively, I investigate the issue of genetically modified (GM) food. I implement a pre-registered survey experiment manipulating both misinformation and corrective information treatments within a panel study based on a probability sample of the general population in Germany. Results show, first, that an emotionally framed piece of misinformation can increase moralization, replicating previous findings. Second, a neutral correction to the misinformation is only effective if the misnformation does not include a strong emotional frame, whereas a correction embedded in a counterframe is effective for both low and high emotional misinformation. In a broader perspective, these results inform further studies and policy interventions on other issues where disinformation loads on strong emotions, ranging from social policy over immigration to health interventions such as vaccinations. \vspace{.8cm}
# abstract-template: Everyone agrees that this issue is really important. But we do not know much about this specific question, although it matters a great deal, for these reasons. We approach the problem from this perspective. Our research design focuses on these cases and relies on these data, which we analyze using this method. Results show what we have learned about the question. They have these broader implications.
